Structured Versus Decorated Cospans

### 2020-11-17 cospans.tex
fixed a typo

This double category is in fact fibrant and so a result of Shulman \cite{Shul} allows us to extract from this symmetric monoidal double category $F \mathbb{C}$sp a symmetric monoidal bicategory $F \mathbf{Csp}$ as the `horizontal bicategory' of the double category $F \mathbb{C}$sp. This bicategory $F \mathbf{Csp}$ has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$,}
\item{morphisms as the horizontal 1-cells of $F\mathbb{C}$sp, which are $F$-decorated cospans of $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$d \in F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of $F$-decorated cospans in $\A$ with the same feet together with a natrual transformation.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b'$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b')$};
\node (F) at (3.65,0) {$\Swarrow \tau$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i'$} (B')
(C) edge node[below]{$o'$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d'$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}
This bicategory $F\mathbf{Csp}$ is superior to the original bicategory $F$Cospan$(A)$ constructed by the second author \cite{Courser} in that it has the 2-morphisms that one would expect such a bicategory to have, such as (\ref{no2morphism}).

@@

Lastly, we require two globular 2-morphisms $\chi \colon (M_1 \otimes N_1) \odot (M_2 \otimes N_2) \to (M_1 \odot M_2) \otimes (N_1 \odot N_2)$ and $\mu \colon U_{a \otimes b} \to U_a \otimes U_b$, the first of which relates the double functors $\otimes$ and $\odot$ and the second of which relates the double functors $\otimes$ and $U$. The latter of these is easy to see: given two objects $a$ and $b$ in $F\lCsp_0=\A$, $\mu$ is given by the identity map of cospans in $\A$:

@@

\begin{proof}
This was shown by Moeller and the third author \cite[Theorems~3.13\&4.1]{MV}. In summary, for a cocartesian base we have correspondences
\begin{gather}
\textrm{lax monoidal pseudofunctors }F\maps(\A,+)\to(\Cat,\times) \notag\\
\Updownarrow \notag\\
\textrm{monoidal opfibrations }U\maps(\X,\otimes)\to(\A,+) \label{monGroth}\\
\Updownarrow \notag\\
\textrm{pseudofunctors }F\maps\A\to\bicat{MonCat} \notag
\end{gather}
The second equivalence was also earlier observed by Shulman \cite{Shulman2008}. In more detail, if $(\phi,\phi_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{sec:2cats}, each fibre category $\X_a=F(a)$ obtains a monoidal structure via
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\phi_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\mathbf{1}\xrightarrow{\phi_0}F(0)\xrightarrow{F(!)}F(a)
\end{equation}
where $\nabla$ is the fold map. 
\end{proof}

@@

Finally, if the original pseudofunctor is in fact lax monoidal $(F,\phi,\phi_0)$, the monoidal Grothendieck construction of \cref{lem:MonGroth} in the cocartesian case expresses $\bot_a$ as follows by \cref{eq:explicitstructure1}:
\begin{equation}\label{eq:initialinfibre}
\mathbf{1}\xrightarrow{\phi_0}F(0_\A)\xrightarrow{F(u_a)}F(a) 
\end{equation}
\end{rmk}

@@

By \cref{cor:fcocMonGroth}, when $F$ as an ordinary pseudofunctor factors through $\Rex$, the Gro\-the\-ndieck construction gives rise to a finitely cocomplete $\inta F$ such that the corresponding opfibration $U\maps\inta{F}\to\A$ preserves all finite colimits. Moreover, by \cref{prop:opfibtolari}, $U$ has a left adjoint $L\maps\A\to\inta F$ with $UL=1_\A$. Diagrammatically,
\begin{displaymath}
 F\maps\A\to\Cat\quad\mapsto\quad\begin{tikzcd}[baseline=.3]\inta F\ar[d,"U"'] \\ \A \end{tikzcd}\quad\mapsto\quad\begin{tikzcd}\A\ar[r,bend left,pos=.55,"L"]\ar[r,phantom,"\bot"description] & \inta F\ar[l,bend left,pos=.45,"U"]\end{tikzcd}
\end{displaymath}
describes the construction of the adjunction from the original $F$. Explicitly, the left adjoint picks the initial object in the finitely cocomplete fibre $L(a)=(a,\bot_a)$, also expressed as $F(u_a)\circ\phi_0(*)$ according to \cref{eq:initialinfibre} for $(\phi,\phi_0)$ the monoidal structure of $F$.

@@

A \define{monoidal} bicategory $\bicat{A}$ comes with a pseudofunctor $\otimes\maps\bicat{A}\times\bicat{A}\to\bicat{A}$ and a unit object $I$ that are associative and unital up to coherent equivalence. A \define{braided} monoidal bicategory comes with a pseudonatural equivalence $\beta_{a,b}\maps a\ot b\to b\ot a$ and appropriate invertible modifications; it is \define{sylleptic} if there is an appropriate invertible modification $1_{a\ot b}\Rrightarrow\beta_{b,a}\circ\beta_{a,b}$ and it is \define{symmetric} if one further axiom holds. A \define{lax monoidal} pseudofunctor between monoidal bicategories $F\maps\bicat{A}\to\bicat{B}$ is a pseudofunctor equipped with pseudonatural transformations with components $\phi_{a,b}\maps Fa\otimes Fb\to F(a\otimes b)$ and $\phi_0\maps I\to FI$ along with coherent invertible modifications for associativity and unitality. This is commonly also called `weak monoidal' pseudofunctor. A \define{braided lax monoidal} pseudofunctor between braided monoidal bicategories comes with an appropriate invertible modification with components $F(\beta_{a,b})\circ \phi_{a,b}\cong\phi_{b,a}\circ \beta_{Fa,Fb}$; a \define{sylleptic lax monoidal} pseudofunctor satisfies one extra condition and a \define{symmetric lax monoidal} pseudofunctor between symmetric monoidal bicategories is just a sylleptic one. RIGHT???

@@

Before formally defining a `double category', it is helpful to have the following picture in mind. A double category has 2-morphisms shaped like:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
We call $A, B, C$ and $D$ \textbf{objects}, $f$ and $g$ \textbf{vertical 1-morphisms}, $M$ and $N$ \textbf{horizontal 1-cells} and $a$ a \textbf{2-morphism}. Note that a vertical 1-morphism is a morphism between objects and a 2-morphism is a morphism between horizontal 1-cells. %We will denote both kinds of morphisms and horizontal 1-cells as a single arrow, namely `$\to$'. 
We follow the notation of Shulman \cite{Shulman2008} with the following definitions.

\begin{defn}\label{defn:double_category}
A \textbf{pseudo double category} $\lD$, or $\textbf{double category}$ for short, consists of a category of objects $\lD_0$ and a category of arrows $\lD_1$ with the following functors
\begin{center}
$U\maps \lD_0 \to \lD_1$\\
$S,T \maps \lD_1 \rightrightarrows \lD_0$\\
$\odot \maps \lD_1 \times_{\lD_0} \lD_1 \to \lD_1$ (where the pullback is taken over $\lD_1 \xrightarrow[]{T} \lD_0 \xleftarrow[]{S} \lD_1$) \\
\end{center}
 such that \\
\begin{center}
$S(U_{A})=A=T(U_{A})$\\
$S(M \odot N)=SN$\\
$T(M \odot N)=TM$\\
\end{center}
equipped with natural isomorphisms
\begin{center}

@@

\end{center}
such that $S(\alpha), S(\lambda), S(\rho), T(\alpha), T(\lambda)$ and $T(\rho)$ are all identities and that the coherence axioms of a monoidal category are satisfied. Following the notation of Shulman, objects of $\lD_0$ are called $\textbf{0-cells}$ and morphisms of $\lD_0$ are called $\textbf{vertical 1-morphisms}$. Objects of $\lD_1$ are called $\textbf{horizontal 1-cells}$ and morphisms of $\lD_1$ are called $\textbf{2-morphisms}$. The morphisms of $\lD_1$, which are vertical 1-morphisms, will be denoted $f \maps A \to C$ and we denote a 1-cell $M$ with $S(M)=A,T(M)=B$ by $M \maps A \to B$. Then a 2-morphism $a \maps M \to N$ of $\lD_1$ with $S(a)=f,T(a)=g$ would look like:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
\end{defn}
A 2-morphism where $f$ and $g$ are identities is called a \textbf{globular 2-morphism}.

%The key difference between a `strict' double category and a pseudo double category is that in a pseudo double category, horizontal composition is associative and unital only up to natural isomorphism. Equivalently, as a double category can be viewed as a category internal to $\Cat$, we can view a pseudo double category as a category `weakly' internal to $\Cat$. We will sometimes omit the word pseudo and simply say double category.

%\begin{defn}
%Let $\lD$ be a pseudo double category. Then the $\textbf{horizontal bicategory}$ of $\lD$, which we denote as $H(\lD)$, is the bicategory consisting of objects of $\lD$, morphisms that are horizontal 1-cells of $\lD$ and 2-morphisms that are globular 2-morphisms of $\lD$.
%\end{defn}

@@

%Following the notation of Shulman \cite{Shulman2010}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:
%\[
%  \xymatrix@-.5pc{
%   A \ar[r]|{|}^{M}  \ar[d]_f \ar@{}[dr]|{\Downarrow a}&
%    B\ar[d]^g\\
%    C \ar[r]|{|}_N & D
%  }
%\]
%We call $M$ and $N$ the \define{horizontal source and target} of the 2-morphism $a$, respectively, and likewise we call $f$ and $g$ the \define{vertical source and target} of the 2-morphism $a$, respectively. Thus $_f \mathbb{A}_g(M,N)$ denotes the set of 2-morphisms in $\mathbb{A}$ with horizontal source and target $M$ and $N$ and vertical source and target $f$ and $g$.
\begin{defn}\label{def:fullfaithful}
A (possibly lax or oplax) double functor $\mathbb{F} \maps \mathbb{A} \to \mathbb{X}$ is \define{full} (respectively, \define{faithful}) if $\mathbb{F}_0 \maps \mathbb{A}_0 \to \mathbb{X}_0$ is full (respectively, faithful) and each map $$\mathbb{F}_1 \maps _f \mathbb{A}_g(M,N) \to _{\mathbb{F}(f)} \mathbb{X}_{\mathbb{F}(g)}(\mathbb{F}(M),\mathbb{F}(N))$$ is surjective (respectively, injective).
\end{defn}
\begin{defn}\label{def:essentiallysurj}
A (possibly lax or oplax) double functor $\mathbb{F} \maps \mathbb{A} \to \mathbb{X}$ is \define{essentially surjective} if we can simultaneously make the following choices:
\begin{enumerate}
\item{For each object $x \in \mathbb{X}$, we can find an object $a \in \mathbb{A}$ together with a vertical 1-isomorphism $\alpha_x \maps \mathbb{F}(a) \to x$, and}
\item{For each horizontal 1-cell $N \maps x_1 \tobar x_2$  of $\mathbb{X}$, we can find a horizontal 1-cell $M \maps a_1 \tobar a_2$ of $\mathbb{A}$ and a 2-isomorphism $a_{N}$ of $\mathbb{X}$ as in the following diagram:
\[
  \xymatrix@-.5pc{
    \mathbb{F}(a_1) \ar[r]|{|}^{\mathbb{F}(M)}  \ar[d]_{\alpha_{x_1}} \ar@{}[dr]|{\Downarrow a_N}&
    \mathbb{F}(a_1) \ar[d]^{\alpha_{x_2}}\\
    x_1 \ar[r]|{|}_N & x_2
  }
\]
}
\end{enumerate}
\end{defn}

Formally, two strong double functors between two double categories form a \emph{double equivalence} if ...
%{\chris their two-way composite is isomorphic to the identity, better expressed, possibly in a definition environment}.   Say this or leave it out entirely!!!
Similarly to ordinary equivalence of categories, \cite[Theorem 7.8]{Shulman2010} allows us to use the following equivalent characterization.

@@

\item $\lD_0$ and $\lD_1$ are both monoidal categories.
\item If $I$ is the monoidal unit of $\lD_0$, then $U_I$ is the
  monoidal unit of $\lD_1$.

>>>>

This double category is in fact fibrant and so a result of Shulman \cite{Shul} allows us to extract from this symmetric monoidal double category $F \mathbb{C}$sp a symmetric monoidal bicategory $F \mathbf{Csp} \mapseqq H(F \lCsp)$ as the `horizontal edge bicategory' of the double category $F \mathbb{C}$sp. This bicategory $F \mathbf{Csp}$ has:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$,}
\item{morphisms as the horizontal 1-cells of $F\mathbb{C}$sp, which are $F$-decorated cospans of $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$d \in F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of $F$-decorated cospans in $\A$ with the same feet together with a natrual transformation.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b'$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b')$};
\node (F) at (3.65,0) {$\Swarrow \tau$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i'$} (B')
(C) edge node[below]{$o'$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d'$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
}
\end{enumerate}
This bicategory $F\mathbf{Csp}$ is superior to the original bicategory $F$Cospan$(A)$ constructed by the second author \cite{Courser} in that it has the 2-morphisms that one would expect such a bicategory to have, such as (\ref{no2morphism}).

@@

Lastly, we require two globular 2-morphisms $\chi \colon (M_1 \otimes N_1) \odot (M_2 \otimes N_2) \to (M_1 \odot M_2) \otimes (N_1 \odot N_2)$ and $\mu \colon U_{a \otimes b} \to U_a \otimes U_b$, the first of which relates the double functor $\otimes$ with the functor $\odot$ and the second of which relates the double functor $\otimes$ with the functor $U$. The latter of these is easy to see: given two objects $a$ and $b$ in $F\lCsp_0=\A$, $\mu$ is given by the identity map of cospans in $\A$:

@@

\begin{proof}
This was shown by Moeller and the third author \cite[Theorems~3.13\&4.1]{MV}. In summary, for a cocartesian base we have correspondences
\begin{gather}
\textrm{lax monoidal pseudofunctors }F\maps(\A,+)\to(\Cat,\times) \notag\\
\Updownarrow \notag\\
\textrm{monoidal opfibrations }U\maps(\X,\otimes)\to(\A,+) \label{monGroth}\\
\Updownarrow \notag\\
\textrm{pseudofunctors }F\maps\A\to\bicat{MonCat} \notag
\end{gather}
The second equivalence was also earlier observed by Shulman \cite{Shulman2008}. In more detail, if $(\mu,\mu_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{sec:2cats}, each fibre category $\X_a=F(a)$ obtains a monoidal structure via
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\mu_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\mathbf{1}\xrightarrow{\mu_0}F(0)\xrightarrow{F(!)}F(a)
\end{equation}
where $\nabla$ is the fold map. 
\end{proof}

@@

Finally, if the original pseudofunctor is in fact lax monoidal $(F,\mu,\mu_0)$, the monoidal Grothendieck construction of \cref{lem:MonGroth} in the cocartesian case expresses $\bot_a$ as follows by \cref{eq:explicitstructure1}:
\begin{equation}\label{eq:initialinfibre}
\mathbf{1}\xrightarrow{\mu_0}F(0_\A)\xrightarrow{F(u_a)}F(a) 
\end{equation}
\end{rmk}

@@

By \cref{cor:fcocMonGroth}, when $F$ as an ordinary pseudofunctor factors through $\Rex$, the Gro\-the\-ndieck construction gives rise to a finitely cocomplete $\inta F$ such that the corresponding opfibration $U\maps\inta{F}\to\A$ preserves all finite colimits. Moreover, by \cref{prop:opfibtolari}, $U$ has a left adjoint $L\maps\A\to\inta F$ with $UL=1_\A$. Diagrammatically,
\begin{displaymath}
 F\maps\A\to\Cat\quad\mapsto\quad\begin{tikzcd}[baseline=.3]\inta F\ar[d,"U"'] \\ \A \end{tikzcd}\quad\mapsto\quad\begin{tikzcd}\A\ar[r,bend left,pos=.55,"L"]\ar[r,phantom,"\bot"description] & \inta F\ar[l,bend left,pos=.45,"U"]\end{tikzcd}
\end{displaymath}
describes the construction of the adjunction from the original $F$. Explicitly, the left adjoint picks the initial object in the finitely cocomplete fibre $L(a)=(a,\bot_a)$, also expressed as $F(u_a)\circ\mu_0(*)$ according to \cref{eq:initialinfibre} for $(\mu,\mu_0)$ the monoidal structure of $F$.

@@

A \define{monoidal} bicategory $\bicat{A}$ comes with a pseudofunctor $\otimes\maps\bicat{A}\times\bicat{A}\to\bicat{A}$ and a unit object $I$ that are associative and unital up to coherent equivalence. A \define{braided} monoidal bicategory comes with a pseudonatural equivalence $\beta_{a,b}\maps a\ot b\to b\ot a$ and appropriate invertible modifications; it is \define{sylleptic} if there is an appropriate invertible modification $1_{a\ot b}\Rrightarrow\beta_{b,a}\circ\beta_{a,b}$ and it is \define{symmetric} if one further axiom holds. A \define{lax monoidal} pseudofunctor between monoidal bicategories $F\maps\bicat{A}\to\bicat{B}$ is a pseudofunctor equipped with pseudonatural transformations with components $\mu_{a,b}\maps Fa\otimes Fb\to F(a\otimes b)$ and $\mu_0\maps I\to FI$ along with coherent invertible modifications for associativity and unitality. This is commonly also called `weak monoidal' pseudofunctor. A \define{braided lax monoidal} pseudofunctor between braided monoidal bicategories comes with an appropriate invertible modification with components $F(\beta_{a,b})\circ \mu_{a,b}\cong\mu_{b,a}\circ \beta_{Fa,Fb}$; a \define{sylleptic lax monoidal} pseudofunctor satisfies one extra condition and a \define{symmetric lax monoidal} pseudofunctor between symmetric monoidal bicategories is just a sylleptic one. RIGHT???

@@

Before formally defining `pseudo double category', it is helpful to have the following picture in mind. A pseudo double category has 2-morphisms shaped like:

\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]

We call $A, B, C$ and $D$ \textbf{objects} or \textbf{0-cells}, $f$ and $g$ \textbf{vertical 1-morphisms}, $M$ and $N$ \textbf{horizontal 1-cells} and $a$ a \textbf{2-morphism}. Note that a vertical 1-morphism is a morphism between 0-cells and a 2-morphism is a morphism between horizontal 1-cells. We will denote both kinds of morphisms and horizontal 1-cells as a single arrow, namely `$\to$'. We follow the notation of Shulman \cite{Shulman2008} with the following definitions.

\begin{defn}\label{defn:double_category}
A \textbf{pseudo double category} $\lD$, or $\textbf{double category}$ for short, consists of a category of objects $\mathbf{D_{0}}$ and a category of arrows $\mathbf{D_1}$ with the following functors
\begin{center}
$U\maps \mathbf{D_{0}} \to \mathbf{D_1}$\\
$S,T \maps \mathbf{D_1} \rightrightarrows \mathbf{D_{0}}$\\
$\odot \maps \mathbf{D_1} \times_{\mathbf{D_{0}}} \mathbf{D_1} \to \mathbf{D_1}$ (where the pullback is taken over $\mathbf{D_1} \xrightarrow[]{T} \mathbf{D_{0}} \xleftarrow[]{S} \mathbf{D_1}$) \\
\end{center}
 such that \\
\begin{center}
$S(U_{A})=A=T(U_{A})$\\
$S(M \odot N)=SN$\\
$T(M \odot N)=TM$\\
\end{center}
equipped with natural isomorphisms
\begin{center}

@@

\end{center}
such that $S(\alpha), S(\lambda), S(\rho), T(\alpha), T(\lambda)$ and $T(\rho)$ are all identities and that the coherence axioms of a monoidal category are satisfied. Following the notation of Shulman, objects of $\mathbf{D_{0}}$ are called $\textbf{0-cells}$ and morphisms of $\mathbf{D_{0}}$ are called $\textbf{vertical 1-morphisms}$. Objects of $\mathbf{D_1}$ are called $\textbf{horizontal 1-cells}$ and morphisms of $\mathbf{D_1}$ are called $\textbf{2-morphisms}$. The morphisms of $\mathbf{D_{0}}$, which are vertical 1-morphisms, will be denoted $f \maps A \to C$ and we denote a 1-cell $M$ with $S(M)=A,T(M)=B$ by $M \maps A \to B$. Then a 2-morphism $a \maps M \to N$ of $\mathbf{D_1}$ with $S(a)=f,T(a)=g$ would look like:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
\end{defn}

The key difference between a `strict' double category and a pseudo double category is that in a pseudo double category, horizontal composition is associative and unital only up to natural isomorphism. Equivalently, as a double category can be viewed as a category internal to $\Cat$, we can view a pseudo double category as a category `weakly' internal to $\Cat$. We will sometimes omit the word pseudo and simply say double category.

\begin{defn}
A 2-morphism where $f$ and $g$ are identities is called a \textbf{globular 2-morphism}.
\end{defn}

\begin{defn}
Let $\lD$ be a pseudo double category. Then the $\textbf{horizontal bicategory}$ of $\lD$, which we denote as $H(\lD)$, is the bicategory consisting of objects of $\lD$, morphisms that are horizontal 1-cells of $\lD$ and 2-morphisms that are globular 2-morphisms of $\lD$.
\end{defn}

@@

Following the notation of Shulman \cite{Shulman2010}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:
\[
  \xymatrix@-.5pc{
    A \ar[r]|{|}^{M}  \ar[d]_f \ar@{}[dr]|{\Downarrow a}&
    B\ar[d]^g\\
    C \ar[r]|{|}_N & D
  }
\]
We call $M$ and $N$ the \define{horizontal source and target} of the 2-morphism $a$, respectively, and likewise we call $f$ and $g$ the \define{vertical source and target} of the 2-morphism $a$, respectively. Thus $_f \mathbb{A}_g(M,N)$ denotes the set of 2-morphisms in $\mathbb{A}$ with horizontal source and target $M$ and $N$ and vertical source and target $f$ and $g$.
\begin{defn}\label{def:fullfaithful}
A (possibly lax or oplax) double functor $\mathbb{F} \maps \mathbb{A} \to \mathbb{X}$ is \define{full} (respectively, \define{faithful}) if $\mathbb{F}_0 \maps \mathbb{A}_0 \to \mathbb{X}_0$ is full (respectively, faithful) and each map $$\mathbb{F}_1 \maps _f \mathbb{A}_g(M,N) \to _{\mathbb{F}(f)} \mathbb{X}_{\mathbb{F}(g)}(\mathbb{F}(M),\mathbb{F}(N))$$ is surjective (respectively, injective).
\end{defn}
\begin{defn}\label{def:essentiallysurj}
A (possibly lax or oplax) double functor $\mathbb{F} \maps \mathbb{A} \to \mathbb{X}$ is \define{essentially surjective} if we can simultaneously make the following choices:
\begin{enumerate}
\item{For each object $x \in \mathbb{X}$, we can find an object $a \in \mathbb{A}$ together with a vertical 1-isomorphism $\alpha_x \maps \mathbb{F}(a) \to x$, and}
\item{For each horizontal 1-cell $N \maps x_1 \tobar x_2$  of $\mathbb{X}$, we can find a horizontal 1-cell $M \maps a_1 \tobar a_2$ of $\mathbb{A}$ and a 2-isomorphism $a_{N}$ of $\mathbb{X}$ as in the following diagram:
\[
  \xymatrix@-.5pc{
    \mathbb{F}(a_1) \ar[r]|{|}^{\mathbb{F}(M)}  \ar[d]_{\alpha_{x_1}} \ar@{}[dr]|{\Downarrow a_N}&
    \mathbb{F}(a_1) \ar[d]^{\alpha_{x_2}}\\
    x_1 \ar[r]|{|}_N & x_2
  }
\]
}
\end{enumerate}
\end{defn}

Formally, two strong double functors between two double categories form a \emph{double equivalence} if ...
%{\chris their two-way composite is isomorphic to the identity, better expressed, possibly in a definition environment}. 
Similarly to ordinary equivalence of categories, [Shulman,7.8] allows us to use the following equivalent characterization.

@@

\item $\mathbf{D_{0}}$ and $\mathbf{D_1}$ are both monoidal categories.
\item If $I$ is the monoidal unit of $\mathbf{D_{0}}$, then $U_I$ is the
  monoidal unit of $\mathbf{D_1}$.
